### 2022-02-22 manuscript/submission_review_2.tex
revisao 2

Large-scale data processing with tens of thousands of data is a reality in all areas of geophysics including the geophysical potential fields. 
The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
The earliest techniques of potential-field data processing  were developed in the space domain.
For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
used in a  graphical convolution with the magnetic data.
However, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
\cite{Dean1958} pointed out that the operations of second order derivatives, analytic continuation, smoothing, the removing of residuals or regionals, and others for processing  potential-field data are
similar to the electric filter circuits in Fourier domain.
\cite{Dean1958} was the first to develop the theory of linear filters in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
for some theoretical geophysical operations (e.g., derivatives and upward and downward continuations).
\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in Fourier domain.

An approach for processing potential-field data in space domain is the equivalent-layer technique.
The equations deductions of the equivalent layer as a solution of the Laplace's equation in the region above the source was first presented by \cite{kellogg1929} and detailed explanations can also be found in \cite{blakely1996}. 
Although the equivalent-layer technique  has been known since the 1960s in geophysical literature \citep{danes1961structure,bott1967solution,dampney1969}, its use has become feasible only recently 
because of the advances in computational power.
In magnetic data procesing, some authors explored this technique for calculating the first and second vertical derivatives of the fields \citep{emilia1973}, reduction to the pole \citep{silva1986,oliveirajr-etal2013,li2014using}, upward/downward continuations \citep{hansen-miyazaki1984,li-oldenburg2010} and total magnetic induction vector components calculation \citep{sun2019constrained}.

@@

In this work, the convolutional equivalent layer using the block-Toeplitz Toeplitz-block idea, presented in \cite{takahashi2020convolutional}, will be used to solve the linear system required to estimate the physical property that produces a magnetic field on regular grids. 
Here, we achieve very fast solutions using a conjugate gradient algorithm combined with the fast Fourier transform. We present a novel method of exploring the symmetric structures of the second derivatives of the inverse of the distance contained in the magnetic kernel, to keep the memory RAM usage to the minimal by using only one equivalent source to carry the calculations of the forward problem. We also show tests of the magnetic convolutional equivalent layer when irregular grids are used. The convergence of the conjugate gradient maintains in an acceptable level even using irregular grids. 
Our results show the good performance of our method in producing fast and robust solutions for processing large amounts of magnetic data using the equivalent layer technique.

@@

%======================================================================================
\subsection{Classical equivalent layer for magnetic data}
%======================================================================================

Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $,
whose $i$th element is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
aranged in a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
pointing to north, east and down, respectively.
We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
values of a harmonic function. Besides, we consider that the main geomagnetic field 
direction is constant at the limited study area and can be defined by the unit vector
\begin{equation}
	\hat{\mathbf{F}} = \begin{bmatrix}
		F_x \\
		F_y \\
		F_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I_{0}) \, \cos(D_{0}) \\
		\cos(I_{0}) \, \sin(D_{0}) \\
		\sin(I_{0})
	\end{bmatrix} \: ,
	\label{eq:unit_vector_F}
\end{equation}
with constant inclination $I_{0}$ and declination $D_{0}$.
In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly
\begin{equation}
	\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
	\label{eq:integral-sum_mag}
\end{equation}
which describes the magnetic induction exerted at the observation point,
by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
\begin{equation}
	a_{ij}
	= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
	\label{eq:aij_mag}
\end{equation}
the unit vector
\begin{equation}
	\hat{\mathbf{u}} = \begin{bmatrix}
		u_x \\
		u_y \\
		u_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I) \, \cos(D) \\
		\cos(I) \, \sin(D) \\
		\sin(I)
	\end{bmatrix} \: ,
	\label{eq:u_hat}
\end{equation}
defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
\begin{equation}
	\mathbf{H}_{ij} = \begin{bmatrix}
		h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
		h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
		h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
	\end{bmatrix} \: ,
	\label{eq:Hij}
\end{equation}
where 
\begin{equation}
	h^{\alpha\beta}_{ij} = 
	\begin{cases}
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
	\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
	\label{eq:hij_alpha_beta}
\end{equation}
are the second derivatives of the inverse distance function
\begin{equation}
	\frac{1}{r_{ij}} = 
	\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
			\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
	\label{eq:1_rij}
\end{equation}
with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
	\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
	\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
	\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
	\mathbf{A}^{\top} \mathbf{d}^{o} \: .
	\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor $\mathbf{G}$ of the positive-definite matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
systems \citep[][ p. 262]{golub-vanloan2013}:
\begin{equation}
	\begin{split}
		\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
		\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
	\end{split} \: ,
	\label{eq:classical-method}
\end{equation}
where $\mathbf{w}$ is a temporarily variable.
This approach to estimating the parameter vector will be 
referenced throughout this work as the \textit{classical method}.
The computational cost associated with the classical method can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

To access the structure of the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}), let us first rewrite its elements 
$a_{ij}$ (equation \ref{eq:aij_mag}) in the following way:
\begin{equation}
	\begin{split}
		a_{ij} = a^{xx}_{ij} + a^{xy}_{ij} + a^{xz}_{ij} + a^{yy}_{ij} + a^{yz}_{ij} + a^{zz}_{ij} \: ,
	\end{split}
	\label{eq:aij_mag_expand}
\end{equation}
where
\begin{equation}
	a^{\alpha\beta}_{ij} = 
	\begin{cases}
		c_{m} \, \frac{\mu_{0}}{4\pi} 
		\left( F_{\alpha} u_{\beta} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha = \beta \\
		c_{m} \, \frac{\mu_{0}}{4\pi} 
		\left( F_{\alpha} u_{\beta} + F_{\beta} u_{\alpha} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha \ne \beta \\
	\end{cases}
	\: , \quad \alpha, \beta = x, y, z \: ,
	\label{eq:aij_alpha_beta}
\end{equation}
are defined by the elements of $\hat{\mathbf{F}}$ 
(equation \ref{eq:unit_vector_F}), $\hat{\mathbf{u}}$ (equation \ref{eq:u_hat}) and 
$\mathbf{H}_{ij}$ (equations \ref{eq:Hij} and \ref{eq:hij_alpha_beta}).
Then, we can rewrite the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}) according to:
\begin{equation}
	\mathbf{A} = \mathbf{A_{xx}} + \mathbf{A_{xy}} + \mathbf{A_{xz}} + 
	\mathbf{A_{yy}} + \mathbf{A_{yz}} + \mathbf{A_{zz}} \: ,
	\label{eq:A_expand}
\end{equation}
where $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ are $N \times M$ matrices with elements 
$ij$ defined by $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}).

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
	x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
	\label{eq:xi}
\end{equation}
and
\begin{equation}
	y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
	\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the start grid coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
	\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
	l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
	\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
	\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
	l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
	\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

@@

	8 - Repeat previous steps until convergence ($\delta$).

@@

Setting a convergence criterion $\delta$ (Algorithm \ref{al:std-cgls-algorithm}) based on the minimum tolerance 
of the residuals is a good option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance to the Euclidean norm of residuals between 
iterations, which would increase algorithm runtime, but with smaller residuals. 
We chose the latter option, as we could achieve better results.

@@

Table \ref{tab:RAM-usage} shows a comparison between the RAM memory storage 
associated with each method. In the classical and standard CGLS methods, the whole 
matrix $\mathbf{A}$ (equation \ref{eq:A_expand}) has to be stored. For example, a dataset with 
$N = 10,000$ observation points has an associated sensitivity matrix $\mathbf{A}$ formed by 
$N^2 = 100,000,000$ elements and takes approximately $763$ Megabytes of memory (8 bytes per element). 
Using the same number of observation points $N = 10,000$, our method requires only 
$1.831$ Megabytes to store the first columns of the BCCB matrices
$\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) and 
$0.6104$ Megabytes to store the complex matrix $\mathbf{L}$ (equation \ref{eq:L}) 
(16 bytes per element). For a bigger dataset with $N = 1,000,000$, the amount of necessary RAM 
goes to $7,629,395$, $183.096$ and $61.035$ Megabytes, respectively.

@@

Finally, Figure \ref{fig:carajas_upward_comparison}a shows the upward-continued magnetic data to a
horizontal plane located at an altitude of $5, \,000$ m using the estimated equivalent layer obtained by applying our
method to the decimated irregular grid. This process took $\approx 2.64$ seconds, showing good results 
without visible errors or border effects. 
Figure \ref{fig:carajas_upward_comparison}b shows the upward-continued magnetic data to a
horizontal plane located at an altitude of $5, \,000$ m using the classical Fourier filtering method to the decimated irregular grid.
This process took $\approx 0.5$ seconds. Altough, no border effect could be observed, this result showed unexpected high values instead of the expected attenuation of the anomalies. We stress that we did not use a padding scheme to expand the data.

>>>>

Large-scale data processing with tens of thousands of data is a reality in all areas of geophysics including the geophysical potential fields. 
The potential-field data processing includes convolution integrals which can be solved either in the space or Fourier domains.
The earliest techniques of potential-field data processing  were developed in the space domain.
For example, \cite{Peters1949} accomplished, in the space domain, the second and fourth derivatives of magnetic data and the upward- and downward-continuations of magnetic data by deriving coefficients that are
used in a  graphical convolution with the magnetic data.
However, the techniques for  processing potential-field data in space-domain were soon substituted by the Fourier-domain techniques. 
\cite{Dean1958} pointed out that the operations of second order derivatives, analytic continuation, smoothing, the removing of residuals or regionals, and others for processing  potential-field data are
similar to the electric filter circuits in Fourier domain.
\cite{Dean1958} was the first to develop the theory of linear filters in Fourier domain for gravity and magnetic processing and to present filters in Fourier domain \citep[][ see Table I, p 113]{Dean1958} 
for some theoretical geophysical operations (e.g., derivatives and upward- and downward-continuations).
\cite{GUNN1975} presented a comprehensive analysis of processing potential-field data in Fourier domain.

An approach for processing potential-field data in space domain is the equivalent-layer technique.
The deductions of the equivalent layer equations as a solution of the Laplace's equation in the region above the source was first presented by \cite{kellogg1929} and detailed explanations can also be found in \cite{blakely1996}. 
Although the equivalent-layer technique  has been known since the 1960s in geophysical literature \citep{danes1961structure,bott1967solution,dampney1969}, its use has become feasible only recently 
because of the advances in computational power.
In magnetic data procesing, some authors explored this technique for calculating the first and second vertical derivatives of the fields \citep{emilia1973}, reduction to the pole \citep{silva1986,oliveirajr-etal2013,li2014using}, upward/downward continuations \citep{hansen-miyazaki1984,li-oldenburg2010} and total magnetic induction vector components calculation \citep{sun2019constrained}.

@@

In this work, the convolutional equivalent layer using the block-Toeplitz Toeplitz-block idea, presented in \cite{takahashi2020convolutional}, will be used to solve the linear system required to estimate the physical property that produces a magnetic field on regular grids. 
Here, we achieve very fast solutions using a conjugate gradient algorithm combined with the fast Fourier transform. We present a novel method of exploring the symmetric structures of the second order derivatives of the inverse of the distance contained in the magnetic kernel, to keep the computer memory usage to the minimal by using only one equivalent source to carry the calculations of the forward problem. We also show tests of the magnetic convolutional equivalent layer when irregular grids are used. The convergence of the conjugate gradient maintains in an acceptable level even using irregular grids. 
Our results show a good performance of our method in producing fast and robust solutions for processing large amounts of magnetic data using the equivalent layer technique.

@@

%======================================================================================
\subsection{Classical equivalent layer for magnetic data}
%======================================================================================
Let $\mathbf{d}^{o}$ be a $N \times 1$ observed data vector where, $d^{o}_{i}$ $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $,
is the total-field anomaly  produced by arbitrarily magnetized sources at the $i$th position,
aranged in a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
pointing to north, east and down, respectively.
%Let $\mathbf{d}^{o}$ be the $N \times 1$ observed data vector $(x_{i}, y_{i}, z_{i})$, $i =  1, \dots, N $,
%whose $i$th element is the total-field anomaly $d^{o}_{i}$ produced by arbitrarily magnetized sources
%aranged in a right-handed Cartesian coordinate system with $x$-, $y$- and $z$-axis 
%pointing to north, east and down, respectively.
We consider that the total-field anomaly data $d^{o}_{i}$ represent the discrete
values of a harmonic function. Besides, we consider that the main geomagnetic field 
direction is constant at the limited study area and can be defined by the unit vector
\begin{equation}
	\hat{\mathbf{F}} = \begin{bmatrix}
		F_x \\
		F_y \\
		F_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I_{0}) \, \cos(D_{0}) \\
		\cos(I_{0}) \, \sin(D_{0}) \\
		\sin(I_{0})
	\end{bmatrix} \: ,
	\label{eq:unit_vector_F}
\end{equation}
with constant inclination $I_{0}$ and declination $D_{0}$.
In this case, $d^{o}_{i}$ can be approximated by the predicted total-field anomaly \citep{blakely1996}
\begin{equation}
	\Delta T_{i} = \sum_{j=1}^{M} \, p_{j} a_{ij} \: ,
	\label{eq:integral-sum_mag}
\end{equation}
which describes the magnetic induction exerted at the observation point,
by a discrete layer of $M$ dipoles (equivalent sources) defined on the horizontal plane $z = z_{c}$, 
where $p_{j}$ is the magnetic moment intensity (in A~m~$^{2}$)~of the $j$th dipole, 
that has unit volume and is located at the point $(x_{j}, y_{j}, z_{c})$. In equation
\ref{eq:integral-sum_mag}, $a_{ij}$ is the harmonic function
\begin{equation}
	a_{ij}
	= c_{m} \, \frac{\mu_{0}}{4\pi} \, \hat{\mathbf{F}}^{\top} \mathbf{H}_{ij} \: \hat{\mathbf{u}} \: ,
	\label{eq:aij_mag}
\end{equation}
the unit vector
\begin{equation}
	\hat{\mathbf{u}} = \begin{bmatrix}
		u_x \\
		u_y \\
		u_z
	\end{bmatrix} =
	\begin{bmatrix}
		\cos(I) \, \cos(D) \\
		\cos(I) \, \sin(D) \\
		\sin(I)
	\end{bmatrix} \: ,
	\label{eq:u_hat}
\end{equation}
defines the magnetization direction of all dipoles, with constant inclination $I$ and declination $D$,
$\mu_{0} = 4\pi \, 10^{-7}$ H/m is the magnetic constant, $c_{m} = 10^{9}$ is a factor that transforms
the magnetic induction from Tesla (T) to nanotesla (nT) and $\mathbf{H}_{ij}$ is a $3 \times 3$ matrix 
\begin{equation}
	\mathbf{H}_{ij} = \begin{bmatrix}
		h^{xx}_{ij} & h^{xy}_{ij} & h^{xz}_{ij} \\
		h^{xy}_{ij} & h^{yy}_{ij} & h^{yz}_{ij} \\
		h^{xz}_{ij} & h^{yz}_{ij} & h^{zz}_{ij}
	\end{bmatrix} \: ,
	\label{eq:Hij}
\end{equation}
where 
\begin{equation}
	h^{\alpha\beta}_{ij} = 
	\begin{cases}
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: , \quad \alpha = \beta \\
		\frac{3 \left( \alpha_{i} - \alpha_{j} \right) \left( \beta_{i} - \beta_{j} \right)}{r_{ij}^{5}} \: , \quad \alpha \ne \beta
	\end{cases} \: , \quad \alpha, \beta = x, y, z \: ,
	\label{eq:hij_alpha_beta}
\end{equation}
are the second derivatives of the inverse distance function
\begin{equation}
	\frac{1}{r_{ij}} = 
	\frac{1}{\sqrt{\left(x_{i} - x_{j} \right)^{2} + 
			\left(y_{i} - y_{j} \right)^{2} + \left(z_{i} - z_{c} \right)^{2}}}
	\label{eq:1_rij}
\end{equation}
with respect to the coordinates of the observation point $(x_{i}, y_{i}, z_{i})$.

Equation \ref{eq:integral-sum_mag} can be rewritten in matrix notation as follows:
\begin{equation}
	\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
	\label{eq:predicted-data-vector_mag}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is the $N \times 1$ predicted data vector with $i$th element defined
as the predicted total-field anomaly $\Delta T_{i}$ (equation \ref{eq:integral-sum_mag}),
$\mathbf{p}$ is the $M \times 1$ parameter vector whose $j$th element is the magnetic moment intensity
$p_{j}$ of the $j$th dipole and $\mathbf{A}$ is the $N \times M$ sensitivity matrix with element 
$ij$ defined by the harmonic function $a_{ij}$ (equation \ref{eq:aij_mag}).
In the classical equivalent-layer technique, the common approach for 
estimating the parameter vector $\mathbf{p}$ from the observed 
total-field anomaly data $\mathbf{d}^{o}$ is solving the least-squares normal equations
\begin{equation}
	\mathbf{A}^{\top}\mathbf{A} \: \mathbf{p} = 
	\mathbf{A}^{\top} \mathbf{d}^{o} \: .
	\label{eq:normal-equations}
\end{equation}
Equation \ref{eq:normal-equations} is usually solved by first computing the Cholesky 
factor $\mathbf{G}$ of the positive-definite matrix $\mathbf{A}^{\top}\mathbf{A}$ and then using it to solve the linear 
systems \citep[][ p. 262]{golub-vanloan2013}:
\begin{equation}
	\begin{split}
		\mathbf{G} \mathbf{w} &= \mathbf{A}^{\top}\mathbf{d}^{o} \\
		\mathbf{G}^{\top} \tilde{\mathbf{p}} &= \mathbf{w}
	\end{split} \: ,
	\label{eq:classical-method}
\end{equation}
where $\mathbf{w}$ is a temporary variable.
This approach to estimating the parameter vector will be 
referenced throughout this work as the \textit{classical method}.
The computational cost associated with the classical method can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

To access the structure of the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}), let us first rewrite its elements 
$a_{ij}$ (equation \ref{eq:aij_mag}) in the following way:
\begin{equation}
	\begin{split}
		a_{ij} = a^{xx}_{ij} + a^{xy}_{ij} + a^{xz}_{ij} + a^{yy}_{ij} + a^{yz}_{ij} + a^{zz}_{ij} \: ,
	\end{split}
	\label{eq:aij_mag_expand}
\end{equation}
where
\begin{equation}
	a^{\alpha\beta}_{ij} = 
	\begin{cases}
		c_{m} \, \frac{\mu_{0}}{4\pi} 
		\left( F_{\alpha} u_{\beta} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha = \beta \\
		c_{m} \, \frac{\mu_{0}}{4\pi} 
		\left( F_{\alpha} u_{\beta} + F_{\beta} u_{\alpha} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha \ne \beta \\
	\end{cases}
	\: , \quad \alpha, \beta = x, y, z \: ,
	\label{eq:aij_alpha_beta}
\end{equation}
are defined by the elements of $\hat{\mathbf{F}}$ 
(equation \ref{eq:unit_vector_F}), $\hat{\mathbf{u}}$ (equation \ref{eq:u_hat}) and 
$\mathbf{H}_{ij}$ (equations \ref{eq:Hij} and \ref{eq:hij_alpha_beta}).
Then, we can rewrite the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}) as:
\begin{equation}
	\mathbf{A} = \mathbf{A_{xx}} + \mathbf{A_{xy}} + \mathbf{A_{xz}} + 
	\mathbf{A_{yy}} + \mathbf{A_{yz}} + \mathbf{A_{zz}} \: ,
	\label{eq:A_expand}
\end{equation}
where $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ are $N \times M$ matrices with elements 
$ij$ defined by $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}).

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
	x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
	\label{eq:xi}
\end{equation}
and
\begin{equation}
	y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
	\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the lower left corners coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
	\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
	l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
	\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
	\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
	l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
	\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

@@

	8 - Repeat previous steps until convergence (stops if $\delta = |\mathbf{r}_{(it + 1)} - \mathbf{r}_{(it)}| \leq 10^{-3}$).

@@

Setting a convergence criterion $\delta$ (Algorithm \ref{al:std-cgls-algorithm}) based on the minimum tolerance 
of the residuals is a good option to carry out this algorithm efficiently and still obtaining very good results. 
Another possibility is to set an invariance limit to the normalized Euclidean norm of residuals between iterations,
which would increase algorithm runtime, but with smaller residuals. 
We chose the latter option, as we could achieve better results.

@@

Table \ref{tab:RAM-usage} shows a comparison between the computer memory storage 
associated with each method. In the classical and standard CGLS methods, the whole 
matrix $\mathbf{A}$ (equation \ref{eq:A_expand}) has to be stored. For example, a dataset with 
$N = 10,000$ observation points has an associated sensitivity matrix $\mathbf{A}$ formed by 
$N^2 = 100,000,000$ elements and takes approximately $763$ Megabytes of memory (8 bytes per element). 
Using the same number of observation points $N = 10,000$, our method requires only 
$1.831$ Megabytes to store the first columns of the BCCB matrices
$\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) and 
$0.6104$ Megabytes to store the complex matrix $\mathbf{L}$ (equation \ref{eq:L}) 
(16 bytes per element). For a bigger dataset with $N = 1,000,000$, the amount of necessary RAM 
goes to $7,629,395$, $183.096$ and $61.035$ Megabytes, respectively.

@@

Finally, Figure \ref{fig:carajas_upward_comparison}a shows the upward-continued magnetic data to a
horizontal plane located at an altitude of $5, \,000$ m using the estimated equivalent layer obtained by applying our
method to the decimated irregular grid. This process took $\approx 2.64$ seconds, showing good results 
without visible errors or border effects. 
Figure \ref{fig:carajas_upward_comparison}b shows the upward-continued magnetic data to a
horizontal plane located at an altitude of $5, \,000$ m using the classical Fourier filtering method to the decimated irregular grid.
This process took $\approx 0.5$ seconds. Altough, no border effect could be observed, this result showed unexpected high values in some regions, for example, the upper left corner,
which has values near to the same magnitude as the observed data, despite the general attenuation of the anomalies occured as expected. We stress that we did not use a padding scheme to expand the data.
